---
title: "Car Assessment Report"
author: "James Bond"
date: "31 December, 2017"
output: html_document
---
Vehicle Analysis
=======

I have analysed `r nrow(mtcars)` cars systematically. The 
following vehicles have the largest `horsepower`:

```{r, message=FALSE}
library(dplyr)
mtcars %>%
  mutate(name = rownames(.)) %>%
  arrange(desc(hp)) %>%
  select(name, hp) %>%
  head(2)
```

The following graph shows that these cars have impressive
**horsepower** and **1/4 mile time**.

```{r, echo=FALSE, warning=FALSE,fig.width=8, fig.height=4}
library(ggplot2)
mtcars %>%
  mutate(name = rownames(.),
         hp_rank = dense_rank(desc(hp)),
         top3 = hp_rank <= 2) %>%
  ggplot(aes(x=qsec, y=hp)) +
  stat_smooth(method = "lm") +
  geom_point(aes(colour=top3)) +
  geom_text(aes(label=ifelse(top3, name, NA))) +
  labs(x="1/4 Mile Time (Seconds)",
       y="Horsepower",
       colour="Top 2 Horsepower")
```

## Selection Criteria

Essential gadgets are required for Health and Safety reasons:

* Rocket launchers (_front_ and _back_)
* Ejectable seats
* Bulletproof screen

## Results

Despite the vehicle have outstanding performance, they are not
fitted with essential health and safety equipment. As a result,
none of them pass the assessment. 