---
title: "Ch3 - Data Transformation - Filter()"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## FILTER()

### Examples


```{r}
# load the required packages
library(dplyr)
library(nycflights13)
# Inspect the data, noting the type of each variable
flights
# View the flights dataset interactively
View(flights)

# Select all flights on January 1st
# Assigning them to a new variable jan1
jan1 <- filter(flights, month == 1, day == 1)
jan1

# Select all flights from November or December
filter(flights, month == 11 | month == 12)

# A useful shorthand for this problem is x %in% y
# It selects every row where x is one of the values in y
filter(flights, month %in% c(11, 12))

# Let’s select only the flights that weren’t delayed
# (on arrival or departure) by more than two hours:
filter(flights, arr_delay <= 120, dep_delay <= 120)
```

### Exercises


Find all flights in the flights dataset that:

1) Had an arrival delay of two or more hours

```{r}
filter(flights, arr_delay >= 120)
```

###########################################################################################
###########################################################################################
2) Flew to Houston (IAH or HOU)

```{r}
filter(flights, dest == "IAH" | dest == "HOU")

```
 

Alternative solution:

```{r}
filter(flights, dest %in% c("IAH", "HOU"))
```

###########################################################################################
###########################################################################################
3) Were operated by United (UA), American (AA) or Delta (DL).

```{r}
filter(flights, carrier %in% c("AA", "DL", "UA"))
```

###########################################################################################
###########################################################################################
4) Departed in summer (July, August and September)

```{r}
filter(flights, month %in%  c(7, 8, 9))
```

###########################################################################################
###########################################################################################
5) Arrived more than two hours late, but did not leave late.

```{r}
filter(flights, dep_delay <= 0, arr_delay > 120)

##
```

###########################################################################################
###########################################################################################
6) Had a departure delay of at least 1 hour, but made up over 30 minutes in flight.

```{r}
filter(flights, dep_delay >= 60, dep_delay - arr_delay > 30)
```

###########################################################################################
###########################################################################################
7) How many flights have a missing dep_time? What other variables are missing?

```{r}
filter(flights, is.na(dep_time))
```

Notably, the arrival time (arr_time) is also missing for these rows. These seem to be cancelled flights.

