---
title: "Ch3 - Data Transformation - Arrange()"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ARRANGE()

### Examples


```{r}
# load the required packages
library(dplyr)
library(nycflights13)

# Arrange flights by year, then month, then day
arrange(flights, year, month, day)

# Use desc() to reorder by a column in descending order
arrange(flights, desc(arr_delay))

# Missing values are always sorted at the end:
df <- tibble(x = c(5, 2, NA))
arrange(df, x)
arrange(df,desc(x))
```

### Exercises

Sort the flights dataset to find:

1) The fastest (in terms of flight time) flights.



```{r}
arrange(flights, air_time)
```


###########################################################################################
###########################################################################################
2) The fastest (in terms of average speed in mph) flights


We can find these flights by sorting by the result of distance / air_time / 60, where the 60 is to convert the expression to miles per hour since air_time is in minutes.

```{r}
arrange(flights, distance / air_time / 60)
```

###########################################################################################
###########################################################################################

3) The flights that departed the most ahead of time.

```{r}
arrange(flights, dep_delay)
```

###########################################################################################
###########################################################################################
4) The flights that were most delayed on departure.

```{r}
arrange(flights, desc(dep_delay))
```


###########################################################################################
###########################################################################################
5) Which flights travelled the furthest?

```{r}
arrange(flights, desc(distance))
```

