---
title: "Ch3 - Data Transformation - mutate()"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MUTATE()

```{r}
library(dplyr)
library(nycflights13)
data(flights)

# Select several columns only
flights_sml <- select(flights,
                year:day,
                ends_with("delay"),
                distance,
                air_time)

# Use these the smaller data frame derive new columns
mutate(flights_sml,
        gain = arr_delay - dep_delay,
        speed = distance / air_time * 60)

# You can refer to columns that you have just created
mutate(flights_sml,
        gain = arr_delay - dep_delay,
        hours = air_time / 60,
        gain_per_hour = gain / hours)

# Keep only the new variables using transmute()
transmute(flights,
          gain = arr_delay - dep_delay,
          hours = air_time / 60,
          gain_per_hour = gain / hour)


# Modular arithmetic: modulo
7 %% 2

# Modular arithmetic: integer division
7 %/% 2

# Compute hour and minute from dep_time
transmute(flights,
          dep_time,
          hour = dep_time %/% 100,
          minute = dep_time %% 100)

# Example of ranking
y <- c(1, 2, 2, NA, 3, 4)
min_rank(y)
```


############################################################
############################################################

1) Currently dep_time and sched_dep_time are convenient to look atbut hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.


Answer:

```{r}
flights_times <- mutate(flights,
                  dep_time_mins = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
                  sched_dep_time_mins = (sched_dep_time %/% 100 * 60 + sched_dep_time %% 100) %% 1440
                )
```


Explanation:


Let's look at dep_time. Currently it gives times in the form 853 = 08:53, 1438 = 14:38 etc. To get the departure times in the number of minutes, integer divide dep_time by 100 using %/% to get the hours since midnight and multiply by 60 to convert to mins. Then add the remainder of dep_time divided by 100. For example, 1504 represents 15:04 (or 3:04 PM), which is

```{r}
(1504 %/% 100 * 60) + (1504 %% 100)
```

904 mins after midnight

There is one remaining issue. Midnight is represented by 2400, which would correspond to 1440 minutes since midnight, but it should correspond to 0. After converting all the times to minutes after midnight, x %% 1440 will convert 1440 to zero, while keeping all the other times the same


############################################################
############################################################
2) Compare air_time with arr_time - dep_time (you need to convert the last two to units of minutes after midnight, see Q1). What do you expect to see? What problem do you see? Why might this be?

Ans:

```{r}

flights_airtime <- 
  mutate(flights,
         dep_time_min = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
         arr_time_min = (arr_time %/% 100 * 60 + arr_time %% 100) %% 1440,
         air_time_2 = arr_time - dep_time,
         air_time_diff = air_time_2 - air_time)

```

We would expect air_time_diff to be zero for all rows, let's see if it is:

```{r}
# in ascending order:
flights_airtime %>% 
  select(arr_time,  dep_time, arr_time_min, dep_time_min, air_time, air_time_2, air_time_diff ) %>% 
  arrange(air_time_diff) 

# in descending order:
flights_airtime %>% 
  select(arr_time,  dep_time, arr_time_min, dep_time_min, air_time, air_time_2, air_time_diff ) %>% 
  arrange(desc(air_time_diff))
  
```
It's not!

Causes:

- Flights often cross time zones, so the total air time will be off (by multiples of 60)
- Flights sometimes pass midnight, i.e. arr_time < dep_time. This will result in times that are off my 24 hours (1440 mins)
- Reading the docs, it turns out that *air_time* does NOT include taxiing, unlike dep_time and arr_time which do. 


############################################################
############################################################

3) What does 1:3 + 1:10 return and why? How about 1:3 + 1:9?

```{r}
1:3 + 1:10
```
```{r}
1:3 + 1:9
```

