---
title: "training_examples"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 0 
Use your interactive console to print your name using : print(“hello <my name>”)

```{r}
print("Hello Everyone!")
```

# Task 1
Create your own custom vector of any data type.
```{r }
v      <- c(2,3,4,5,1,5,3,8,1)
len    <- length(v)
middle <- v[ceiling(len/2):ceiling(len/2)+1]
1:3 + 1:10
1:3 + 1:9
```

# Task 2
Create your own favourite movies data frame (Your data frame must contain more than one type.)

```{r}
df <- data.frame("number"     = c(1,2,3,4),
                 "alpha"      = c("a","b","c","d"),
                 "phys_constants"= c(2.998e08, 6.02e23, 8.85e-12,5.01e-11 )
)
```
Return a variable (column) of your data frame.

```{r}
print(df$number)
print(df[,1])
```
Can you add another variable (column) to your data frame?
```{r}
df$animals <- c("monkey","giraffe", "elephant", "hippotamus")
print(head(df))
```
Can you return an observation (row)?

```{r}
print(df[1,])
```

#Task 3
Write a new function called “square.it” that returns the square of a number.


```{r}
square.it <- function(x){
  return(x*x)
}
```
Can you apply this to a numerical column in your dataframe?
```{r}
square.it(df$number)

```

#Task 4
Write a for loop that runs through a list of numbers 1:20 and prints only the even numbers. 
```{r}
is.odd <- function(x){
  remainder <- x%%2
  equalOne  <- remainder == 1
  return(equalOne)
}
```
Can you change this loop so it prints just the odd numbers?
```{r}
for(i in 1:20){
  if(is.odd(i)){
    print(i)
  }
}
```
# Task 5
Write apply function for your Task 4 For loop.
```{r}
vec <- 1:20
sapply(vec, function(x){ if(is.odd(x)){print(x)}})
```

(Advanced) Use system.time to find out if apply is faster than for.

```{r}
system.time(for(i in 1:20){
  if(is.odd(i)){
    print(i)
  }
})

vec <- 1:20
system.time(sapply(vec, function(x){ if(is.odd(x)){print(x)}}))
```

# Task 6
library(dplyr)
library(nycflights13)
head(as.data.frame(flights))

# Tidyverse examples
flight_times <- mutate(flights, dep_time_minutes = ( dep_time %/% 100 * 60 +dep_time%%100)%%1440)
